\thetitle

### 2020-07-02 writing/challenges.tex
fix latex

The second big challenge is the set of seemingly incomparable gradual
typing systems. Despite the fact that the natural, transient and optional
approaches are all well-understood as variants of the same minimal operational model
~\cite{gf-icfp-2018, gfd-oopsla-2019}, their actual implementations live on
different linguistic settings. Hence it is not possible to compare them directly 
on the same programs; code needs to be adjusted in often subtle ways to
transfer from one implementation to another. As a first step to overcome the problem, 
we have implemented a transient and optional variant of Typed Racket.  
However,
implementations on top of the same platform are not sufficient for an
apples-to-apples comparison. As we discuss in detail in
section~\ref{sec:landscape}, the three strategies differ in the kind of
blame information they produce. Natural blames one component while
transient blames a set of components and optional none. To account for these
differences, our competent programmer comes with \emp{modes} that represent 
the alternative ways programmers can interpret blame information. For instance, in
transient, one mode  has the competent programmer pick the oldest element of a blame set 
as the next component to add types. After all, this corresponds to the
earliest point in the program's evaluation that can discover the bug. 
Another mode though, does not distinguish between the elements of the
blame set and instead it requires that the competent programmer types 
all blamed components at once. In order to compare different strategies
across their different modes, we normalize the steps the competent
programmer takes to find the bug by the effort of each step, that is the
number of component it has to type. Moreover, besides differences in how
and what they blame, the three gradual typing approaches differ in the type checks they perform. 
To untangle the two, we supplement our experiment with an extra mode. The \emph{vanilla} mode
ignores blame and instead uses information from dynamic type error messages to
select the next component to type. Hence it separates the effect of blame 
from that of the type checks. Finally, to exclude that the experiment
results are not products of pure chance, the \emph{control} mode randomly
selects which part of the program to migrate and serves as the null hypothesis
control for the overall method.

>>>>

The second big challenge is the set of seemingly incomparable gradual
typing systems. Despite the fact that the natural, transient and optional
approaches are all well-understood as variants of the same minimal operational model
~\cite{gf-icfp-2018, gfd-oopsla-2019}, their actual implementations live on
different linguistic settings. Hence it is not possible to compare them directly 
on the same programs; code needs to be adjusted in often subtle ways to
transfer from one implementation to another. As a first step to overcome the problem, 
we have implemented a transient and optional variant of Typed Racket.  
However,
implementations on top of the same platform are not sufficient for an
apples-to-apples comparison. As we discuss in detail in
section~\ref{sec:landscape}, the three strategies differ in the kind of
blame information they produce. Natural blames one component while
transient blames a set of components and optional none. To account for these
differences, our competent programmer comes with \emph{modes} that represent 
the alternative ways programmers can interpret blame information. For instance, in
transient, one mode  has the competent programmer pick the oldest element of a blame set 
as the next component to add types. After all, this corresponds to the
earliest point in the program's evaluation that can discover the bug. 
Another mode though, does not distinguish between the elements of the
blame set and instead it requires that the competent programmer types 
all blamed components at once. In order to compare different strategies
across their different modes, we normalize the steps the competent
programmer takes to find the bug by the effort of each step, that is the
number of component it has to type. Moreover, besides differences in how
and what they blame, the three gradual typing approaches differ in the type checks they perform. 
To untangle the two, we supplement our experiment with an extra mode. The \emph{vanilla} mode
ignores blame and instead uses information from dynamic type error messages to
select the next component to type. Hence it separates the effect of blame 
from that of the type checks. Finally, to exclude that the experiment
results are not products of pure chance, the \emph{control} mode randomly
selects which part of the program to migrate and serves as the null hypothesis
control for the overall method.
